---
title: "Modelling Binomial Outcome Data using ABDs and BMDs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modelling Binomial Outcome Data using ABDs and BMDs}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fitODBOD)
library(ggplot2)
library(reshape2)
library(grid)
library(gridExtra)
library(bbmle)
library(tibble)
```

>IT WOULD BE CLEARLY BENEFICIAL FOR YOU BY USING THE RMD FILES IN THE GITHUB DIRECTORY FOR FURTHER EXPLANATION
OR UNDERSTANDING OF THE R CODE FOR THE RESULTS OBTAINED IN THE VIGNETTES. 

Fitting Alternate Binomial or Binomial Mixture distribution is the most crucial part of while 
handling Binomial Outcome Data. Finding the most suitable distribution which is similar to the 
given data is very important and essential. In order to compare distributions after fitted we can 
choose several measurements. They are namely

* Comparing actual frequency and estimated frequency.
* Using Chi-squared Test statistic and comparing p-values.
* Comparing actual variance with estimated variances of distributions.
* Comparing Negative Log Likelihood values of distributions.
* Comparing AIC values of distributions. 

The functions given below are used to fit respective distributions when BOD and estimated 
parameters are given. 

* `fitBin` - fitting the Binomial distribution.
* `fitTriBin`- fitting the Triangular Binomial distribution.
* `fitBetaBin` - fitting the Beta-Binomial distribution.
* `fitKumBin` - fitting the Kumaraswamy Binomial distribution.
* `fitGHGBB` - fitting the Gaussian Hyper-geometric Generalized Beta-Binomial distribution.
* `fitMcGBB` - fitting the McDonald Generalized Beta-Binomial distribution.
* `fitAddBin` - fitting the Additive Binomial distribution.
* `fitBetaCorrBin` - fitting the Beta Correlated Binomial distribution.
* `fitCOMPBin` - fitting the COM Poisson Binomial distribution.
* `fitCorrBin` - fitting the Correlated Binomial distribution.
* `fitMultiBin` - fitting the Multiplicative Binomial distribution.

## Fitting Alternate Binomial Distributions

All five Alternate Binomial distributions will be fitted to the Alcohol data week 2 and their expected frequencies 
will be plotted with the actual frequency values. This is plot can be used to identify which distribution suits 
best for the Alcohol data week 2. 

```{r plotting Expected frequencies with actual frequency for ABD,include=FALSE, warning=FALSE}
Alcohol_data

BinRanVar<-Alcohol_data$Days
ActFreq<-Alcohol_data$week2

# Fitting Binoial Distribution
BinFreq<-suppressWarnings(fitBin(BinRanVar,ActFreq,print=F)$exp.freq)

# Estimating and fitting Additive Binomial distribution
Para_AddBin<-suppressWarnings(EstMLEAddBin(BinRanVar,ActFreq))
AddBin_p<-Para_AddBin$p
AddBin_alpha<-Para_AddBin$alpha

AddBinFreq<-suppressWarnings(fitAddBin(BinRanVar, ActFreq, AddBin_p, AddBin_alpha, print=F)$exp.freq)

# Estimating and fitting Beta Correlated Binoial Distribution
Para_BetaCorrBin<-suppressWarnings(mle2(EstMLEBetaCorrBin, 
                       data=list(x=BinRanVar, freq=ActFreq),
                       start=list(cov=0.001,a=10,b=10)))
BetaCorrBin_cov<-coef(Para_BetaCorrBin)[1]
BetaCorrBin_a<-coef(Para_BetaCorrBin)[2]
BetaCorrBin_b<-coef(Para_BetaCorrBin)[3]

BetaCorrBinFreq<-suppressWarnings(fitBetaCorrBin(BinRanVar, ActFreq,BetaCorrBin_cov,BetaCorrBin_a,BetaCorrBin_b,print=F)$exp.freq)

# Estimating and fitting COM Poisson Binomial Distribution

Para_COMPBin<-suppressWarnings(mle2(EstMLECOMPBin, 
                       data=list(x=BinRanVar, freq=ActFreq),
                       start=list(v=12.1,p=0.9)))
COMPBin_p<-coef(Para_COMPBin)[1]
COMPBin_v<-coef(Para_COMPBin)[2]

COMPBinFreq<-suppressWarnings(fitCOMPBin(BinRanVar, ActFreq,COMPBin_p,COMPBin_v,print=F)$exp.freq)

# Estimating and fitting Correlated Binomial Distribution

Para_CorrBin<-suppressWarnings(mle2(EstMLECorrBin, 
                       data=list(x=BinRanVar, freq=ActFreq),
                       start=list(cov=0.0021,p=0.19)))
CorrBin_p<-coef(Para_CorrBin)[1]
CorrBin_cov<-coef(Para_CorrBin)[2]

CorrBinFreq<-suppressWarnings(fitCorrBin(BinRanVar, ActFreq,CorrBin_p,CorrBin_cov,print=F)$exp.freq)

# Estimating and fitting Multiplicative Binomial Distribution

Para_MultiBin<-suppressWarnings(mle2(EstMLEMultiBin, 
                       data=list(x=BinRanVar, freq=ActFreq),
                       start=list(theta=21,p=0.19)))
MultiBin_p<-coef(Para_MultiBin)[1]
MultiBin_theta<-coef(Para_MultiBin)[2]

MultiBinFreq<-suppressWarnings(fitMultiBin(BinRanVar, ActFreq,MultiBin_p,MultiBin_theta,print=F)$exp.freq)

# creating dataset for plotting 

ABD_Data<-tibble(
                  w=BinRanVar,
                  x=ActFreq,
                  y=BinFreq,
                  z=AddBinFreq,
                  a=BetaCorrBinFreq,
                  b=COMPBinFreq,
                  c=CorrBinFreq,
                  d=MultiBinFreq
                )
names(ABD_Data)<-c("Bin_RV","Actual_Freq","EstFreq_BinD","EstFreq_AddBinD","EstFreq_BetaCorrBinD","EstFreq_COMPBinD","EstFreq_CorrBinD","EstFreq_MultiBinD" )
```

```{r plotting that dataset for ABD,fig.align='center',fig.height=7,fig.width=9,echo=FALSE}

ABD_freq<-ggplot(ABD_Data)+theme_get()+
                     geom_point(aes(x=ABD_Data$Bin_RV,y=ABD_Data$Actual_Freq),color="red",size=2)+
                     geom_line(aes(x=ABD_Data$Bin_RV,y=ABD_Data$Actual_Freq),color="red",size=1)+
                     geom_point(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_BinD),color="darkblue",size=2)+
                     geom_line(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_BinD),color="darkblue",size=1)+
                     geom_point(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_AddBinD),color="forestgreen",size=3.25)+
                     geom_line(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_AddBinD),color="forestgreen",size=2.25)+
                     geom_point(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_BetaCorrBinD),color="yellow4",size=2)+
                     geom_line(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_BetaCorrBinD),color="yellow4",size=1)+
                     geom_point(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_COMPBinD),color="purple",size=2)+
                     geom_line(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_COMPBinD),color="purple",size=1)+
                     geom_point(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_CorrBinD),color="orangered",size=2)+
                     geom_line(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_CorrBinD),color="orangered",size=1)+
                     geom_point(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_MultiBinD),color="deeppink",size=2)+
                     geom_line(aes(x=ABD_Data$Bin_RV,y=ABD_Data$EstFreq_MultiBinD),color="deeppink",size=1)+
                     ggtitle("Plot of fitted and actual frequencies of Alcohol BOD week2")+
                     xlab("Binomial Random Variables from 0 to 7")+
                     ylab("Frequency Values Estimated and Actual")+
                     scale_y_continuous(breaks=seq(0,120,5))+
                     scale_x_continuous(breaks=seq(0,7,1))+
                     annotate("text",x=0,y=120,label="Actual",color="red")+
                     annotate("text",x=0,y=116,label="Binomial",color="darkblue")+
                     annotate("text",x=0.25,y=112,label="Additive-Binomial",color="forestgreen")+
                     annotate("text",x=0.45,y=108,label="Beta Correlated Binomial",color="yellow4")+
                     annotate("text",x=6.65,y=120,label="COM Poisson Binomial",color="purple")+  
                     annotate("text",x=6.65,y=116,label="Correlated Binomial",color="orangered")+  
                     annotate("text",x=6.65,y=112,label="Multiplicative Binomial",color="deeppink")

  ABD_freq

```

### Conclusion

It is clearly visible that Additive and Correlated Binomial distributions behave very similarly, both 
generate same frequency values. Although they are close to actual frequencies. Comparing Binomial 
distribution generated expected frequencies with actual frequencies will lead to see that there is very 
much difference and Binomial distribution does not suit this Alcohol data of week 2. Multiplicative 
and COM Poisson Binomial distributions show close values to actual frequencies. Finally, the only distribution
left is Beta Correlated Binomial distribution which shows more closeness to actual frequencies. Therefore 
it is clear that most suitable distribution for alcohol data week 2 is Beta Correlated Binomial 
distribution, second choice is Multiplicative and COM Poisson Binomial distributions and final choice is 
Correlated and Additive Binomial distributions.

## Fitting Binoimal Mixture Distributions

In the six BMD distributions except Uniform Binomial distribution others can be used for fitting the 
Alcohol data week 2. Here also as above a plot was generated to compare estimated frequencies 
with actual frequency. 

```{r plotting Expected frequencies with actual frequency for BMD,include=FALSE, warning=FALSE}
Alcohol_data

BinRanVar<-Alcohol_data$Days
ActFreq<-Alcohol_data$week2

# Fitting Binoial Distribution
BinFreq<-suppressWarnings(fitBin(BinRanVar,ActFreq,print=F)$exp.freq)

# Estimating and fitting Triangular Binomial distribution
Para_TriBin<-suppressWarnings(EstMLETriBin(BinRanVar,ActFreq))
TriBin_c<-Para_TriBin$mode


TriBinFreq<-suppressWarnings(fitTriBin(BinRanVar, ActFreq, TriBin_c, print=F)$exp.freq)

# Estimating and fitting Beta Correlated Binoial Distribution
Para_BetaBin<-suppressWarnings(mle2(EstMLEBetaBin, 
                       data=list(x=BinRanVar, freq=ActFreq),
                       start=list(a=10,b=10)))
BetaBin_a<-coef(Para_BetaBin)[1]
BetaBin_b<-coef(Para_BetaBin)[2]

BetaBinFreq<-suppressWarnings(fitBetaBin(BinRanVar, ActFreq,BetaBin_a,BetaBin_b,print=F)$exp.freq)

# Estimating and fitting Kumaraswamy Binomial Distribution

Para_KumBin<-suppressWarnings(mle2(EstMLEKumBin, 
                       data=list(x=BinRanVar, freq=ActFreq),
                       start=list(a=12.1,b=0.9,it=10000)))
KumBin_a<-coef(Para_KumBin)[1]
KumBin_b<-coef(Para_KumBin)[2]
KumBin_it<-coef(Para_KumBin)[3]

KumBinFreq<-suppressWarnings(fitKumBin(BinRanVar, ActFreq,KumBin_a,KumBin_b,KumBin_it*10,print=F)$exp.freq)

# Estimating and fitting GHGBB Distribution

Para_GHGBB<-suppressWarnings(mle2(EstMLEGHGBB, 
                       data=list(x=BinRanVar, freq=ActFreq),
                       start=list(a=0.0021,b=0.19,c=0.3)))
GHGBB_a<-coef(Para_GHGBB)[1]
GHGBB_b<-coef(Para_GHGBB)[2]
GHGBB_c<-coef(Para_GHGBB)[3]

GHGBBFreq<-suppressWarnings(fitGHGBB(BinRanVar, ActFreq,GHGBB_a,GHGBB_b,GHGBB_c,print=F)$exp.freq)

# Estimating and fitting McGBB Distribution

Para_McGBB<-suppressWarnings(mle2(EstMLEMcGBB, 
                       data=list(x=BinRanVar, freq=ActFreq),
                       start=list(a=21,b=0.19,c=0.1)))
McGBB_a<-coef(Para_McGBB)[1]
McGBB_b<-coef(Para_McGBB)[2]
McGBB_c<-coef(Para_McGBB)[3]

McGBBFreq<-suppressWarnings(fitMcGBB(BinRanVar, ActFreq,McGBB_a,McGBB_b,McGBB_c,print=F)$exp.freq)

# creating dataset for plotting 

BMD_Data<-tibble(
                  w=BinRanVar,
                  x=ActFreq,
                  y=BinFreq,
                  z=TriBinFreq,
                  a=BetaBinFreq,
                  b=KumBinFreq,
                  c=GHGBBFreq,
                  d=McGBBFreq
                )
names(BMD_Data)<-c("Bin_RV","Actual_Freq","EstFreq_BinD","EstFreq_TriBinD","EstFreq_BetaBinD","EstFreq_KumBinD","EstFreq_GHGBBD","EstFreq_McGBBD" )
```

```{r plotting that dataset for BMD,fig.align='center',fig.height=7,fig.width=9,echo=FALSE}

BMD_freq<-ggplot(ABD_Data)+theme_get()+
                     geom_point(aes(x=BMD_Data$Bin_RV,y=BMD_Data$Actual_Freq),color="red",size=2)+
                     geom_line(aes(x=BMD_Data$Bin_RV,y=BMD_Data$Actual_Freq),color="red",size=1)+
                     geom_point(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_BinD),color="darkblue",size=2)+
                     geom_line(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_BinD),color="darkblue",size=1)+
                     geom_point(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_TriBinD),color="forestgreen",size=2)+
                     geom_line(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_TriBinD),color="forestgreen",size=1)+
                     geom_point(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_BetaBinD),color="yellow4",size=3.5)+
                     geom_line(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_BetaBinD),color="yellow4",size=2.5)+
                     geom_point(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_KumBinD),color="purple",size=2.75)+
                     geom_line(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_KumBinD),color="purple",size=1.75)+
                     geom_point(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_GHGBBD),color="orangered",size=2)+
                     geom_line(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_GHGBBD),color="orangered",size=1)+
                     geom_point(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_McGBBD),color="deeppink",size=2.25)+
                     geom_line(aes(x=BMD_Data$Bin_RV,y=BMD_Data$EstFreq_McGBBD),color="deeppink",size=1.25)+
                     ggtitle("Plot of fitted and actual frequencies of Alcohol BOD week2")+
                     xlab("Binomial Random Variables from 0 to 7")+
                     ylab("Frequency Values Estimated and Actual")+
                     scale_y_continuous(breaks=seq(0,120,5))+
                     scale_x_continuous(breaks=seq(0,7,1))+
                     annotate("text",x=0,y=120,label="Actual",color="red")+
                     annotate("text",x=0,y=116,label="Binomial",color="darkblue")+
                     annotate("text",x=0.25,y=112,label="Triangular Binomial",color="forestgreen")+
                     annotate("text",x=0.15,y=108,label="Beta-Binomial",color="yellow4")+
                     annotate("text",x=6.5,y=120,label="Kumaraswamy Binomial",color="purple")+  
                     annotate("text",x=6.95,y=116,label="GHGBB",color="orangered")+  
                     annotate("text",x=6.95,y=112,label="McGBB",color="deeppink")

BMD_freq

```

### Conclusion

Estimated frequencies from Triangular distribution are far more better than estimated
frequencies from Binomial distribution. It is clear that Beta-Binomial, Kumaraswamy Binomial
and McDonald Generalized Beta-Binomial distributions all behave equally for the alcohol
data of week 2. Finally,  Gaussian Hyper-geometric Generalized Beta-Binomial distribution is
best suited and generates more accurate frequencies. Therefore, first choice is GHGBB 
distribution and second choice would be Beta-Binomial, Kumaraswamy Binomial and McGBB distributions.
