---
title: "Minimizing the Negative Log Likelihood for BMD and ABD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Minimizing the Negative Log Likelihood for BMD and ABD}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fitODBOD)
library(ggplot2)
library(reshape2)
library(grid)
library(gridExtra)
library(bbmle)
```

>IT WOULD BE CLEARLY BENEFICIAL FOR YOU BY USING THE RMD FILES IN THE GITHUB DIRECTORY FOR FURTHER EXPLANATION
OR UNDERSTANDING OF THE R CODE FOR THE RESULTS OBTAINED IN THE VIGNETTES.

# Negative Log Likelihood Estimation or Maximum Likelihood Estimation Technique

All of the Binomial Mixture and Alternate Binomial distributions in this package use this technique. 
This is very simple, by using the probability mass function we can develop the Likelihood function and then maximize
it or develop the Negative Log Likelihood function and minimize it. Maximizing or minimizing means finding 
appropriate values for the parameters in question which will provide a most maximum Likelihood value or least
minimum Negative Log Likelihood value. 

If we have one parameter to estimate it is possible to use brute force technique. When the number of parameters
need to be estimated increases this is very time consuming. Similarly, if the parameter in estimation has a very
broad domain also it would be difficult. 

In situations with difficulty of such we can use functions such as `optim`,`mle`, `mle2`, `nlm`, `nlminb`. 
`fitODBOD`is using the `mle2` function of the [`bbmle`](https://cran.r-project.org/package=bbmle) 
package. There are situations where user defined functions without the help of `mle2`.

## Functions which does not need `bbmle` package

These situations occurring are very less in comparing to the number of distributions we have in `fitODBOD`.
They genuinely occur when the parameter in need of estimation has a closed domain and that domain is very small.
Here by small it is indicating a scenario of such as in-between zero and one or greater than zero.  

```{r Printing data}
# estimating parameter for p and alpha using Additive Binomial Distribution of Male_children data
Male_Children
```

The distributions in relation are

1. from Alternate Binomial Distribution - Additive Binomial Distribution

There are two parameters to estimate, which are p and alpha. These parameters are unique values, therefore 
estimating them once is enough. Time related to estimation is dependent on the data. 

```{r printing estimated parameters for Additive Binomial Distribution,echo=FALSE}
# estimating parameter for p and alpha using Additive Binomial Distribution
Est_para<-suppressWarnings(EstMLEAddBin(Male_Children$No_of_Males,Male_Children$freq))
cat("Estimated parameter p for Male_children data=",Est_para$p, "\n")
cat("Estimated parameter alpha for Male_children data=",Est_para$alpha,"\n")
```

2. from Binomial Mixture Distribution - Triangular Binomial Distribution

There is only one parameter that needs to be estimated, which is mode or c. This is in-between 
zero and one. 
      
```{r printing estimated parameter for Triangular Binomial Distribution,echo=FALSE}      
# estimating parameter for mode using Triangular Binomial Distribution of Male_children data
Est_para<-suppressWarnings(EstMLETriBin(Male_Children$No_of_Males,Male_Children$freq))
cat("Estimated parameter mode or c for Male_children data=",Est_para$mode, "\n")
```      

## Functions which does need `bbmle` package

By Using the [vignette file](https://cran.r-project.org/package=bbmle) or 
[manual pdf file](https://cran.r-project.org/package=bbmle) we can get to know how the `bbmle` 
package works. Specifically looking into `mle2` function is necessary and also class of `mle2`. 
Notable points related are in point form as 

* coef
* min 
* nobs
* minuslogl
* method

It is preferable to try them and understand the functions capability and process. You can achieve proper results 
by looking at how the inputs should mentioned, because there is a specific format to follow.

Below are instances of using the [`mle2`](https://www.rdocumentation.org/packages/bbmle/versions/1.0.20/topics/mle2) function.

The distributions in relation are 

### from Alternate Binomial distributions 

1. Beta-Correlated Binomial distribution

Here there are three parameters to be estimated which are cov, a and b. It should be mentioned that
cov is a unique value but alpha and beta can take specific combinations.

* cov is in the domain region of -infinity and +infinity.
* a, b are domain in the region of greater than zero.

```{r printing estimated parameters for Beta-Correlated Binomial Distribution, echo=FALSE}
#estimating parameters for Beta-Correlated Binomial Distribution of Male_children data
Est_para<-suppressWarnings(mle2(EstMLEBetaCorrBin,
                           data=list(x=Male_Children$No_of_Males,freq=Male_Children$freq),
                           start=list(cov=0.05,a=10,b=10)))
cat("Estimated parameter covariance for Male_children data=",coef(Est_para)[1],"\n")
cat("Estimated parameter a for Male_children data=",coef(Est_para)[2], "\n")
cat("Estimated parameter b for Male_children data=",coef(Est_para)[3], "\n")
```

2. Correlated Binomial distribution

Two parameters that needs to be estimated which are cov and p. They are unique values for the distribution.

* cov is in the domain region of -infinity and +infinity.
* p is in the domain region of zero and one.

```{r printing estimated parameters for Correlated Binomial Distribution,echo=FALSE}
#estimating parameters for Correlated Binomial Distribution of Male_children data
Est_para<-suppressWarnings(mle2(EstMLECorrBin,
                           data=list(x=Male_Children$No_of_Males,freq=Male_Children$freq),
                           start=list(cov=0.0005,p=0.51)))
cat("Estimated parameter p for Male_children data=",coef(Est_para)[1],"\n")
cat("Estimated parameter covariance for Male_children data=",coef(Est_para)[2], "\n")
```

3. COM Poisson Binomial distribution

Two parameters that needs to be estimated which are p and cov. They are unique values for the distribution.

* cov is in the domain region of -infinity and +infinity.
* p is in the domain region of zero and one.

```{r printing estimated parameters for COM Poisson Binomial Distribution,echo=FALSE}
#estimating parameters for COM Poisson Binomial Distribution of Male_children data
Est_para<-suppressWarnings(mle2(EstMLECOMPBin,
                           data=list(x=Male_Children$No_of_Males,freq=Male_Children$freq),
                           start=list(v=0.0005,p=0.51)))
cat("Estimated parameter p for Male_children data=",coef(Est_para)[1],"\n")
cat("Estimated parameter covariance for Male_children data=",coef(Est_para)[2], "\n")
```

4. Multiplicative Binomial distribution

Two parameters that needs to be estimated which are p and theta. Only p value is unique.

* p is in the domain region of zero and one.
* theta is in the domain region of greater than zero.

```{r printing estimated parameters for Multiplicative Binomial Distribution,echo=FALSE}
#estimating parameters for Multiplicative Binomial Distribution of Male_children data
Est_para<-suppressWarnings(mle2(EstMLEMultiBin,
                           data=list(x=Male_Children$No_of_Males,freq=Male_Children$freq),
                           start=list(theta=15,p=0.51)))
cat("Estimated parameter p for Male_children data=",coef(Est_para)[1],"\n")
cat("Estimated parameter theta for Male_children data=",coef(Est_para)[2], "\n")
```

### from Binomial Mixture distributions

1. Beta-Binomial distribution

Two parameters to be estimated which are a and b. Together they have to represent a unique combination.

* a,b in the domain region of greater than zero.

```{r printing estimated parameters for Beta-Binomial Distribution,echo=FALSE}
#estimating parameters for Beta-Binomial Distribution of Male_children data
Est_para<-suppressWarnings(mle2(EstMLEBetaBin,
                           data=list(x=Male_Children$No_of_Males,freq=Male_Children$freq),
                           start=list(a=15,b=0.51)))
cat("Estimated parameter a for Male_children data=",coef(Est_para)[1],"\n")
cat("Estimated parameter b for Male_children data=",coef(Est_para)[2], "\n")
```

2. Kumaraswamy Binomial distribution

Two parameters to be estimated which are a and b. Together they have to represent a unique combination.

* a,b in the domain region of greater than zero.
* it value or iteration value as a variable replacing the infinity in the summation.

Here the number of iterations has to be estimated as well, which is time consuming. 
therefore I haven't estimated it. 

```{r printing estimated parameters for Kumaraswamy Binomial Distribution, echo=FALSE}
#estimating parameters for Kumaraswamy Binomial Distribution of Male_children data
Est_para<-suppressWarnings(mle2(EstMLEKumBin,
                           data=list(x=Male_Children$No_of_Males,freq=Male_Children$freq),
                           start=list(a=15,b=10,it=5000)))
cat("Estimated parameter a for Male_children data=",coef(Est_para)[1],"\n")
cat("Estimated parameter b for Male_children data=",coef(Est_para)[2], "\n")
cat("Estimated parameter iteration for Male_children data=",coef(Est_para)[3], "\n")
```

3. GHGBB distribution

Three parameters to be estimated which are a,b and c. They have to represent a unique combination as a whole.

* a,b,c in the domain region of greater than zero.

```{r printing estimated parameters for GHGBB Distribution, echo=FALSE}
#estimating parameters for GHGBB Distribution of Male_children data
Est_para<-suppressWarnings(mle2(EstMLEGHGBB,
                           data=list(x=Male_Children$No_of_Males,freq=Male_Children$freq),
                           start=list(a=15,b=10,c=20)))
cat("Estimated parameter a for Male_children data=",coef(Est_para)[1],"\n")
cat("Estimated parameter b for Male_children data=",coef(Est_para)[2], "\n")
cat("Estimated parameter c for Male_children data=",coef(Est_para)[3], "\n")
```

4. McGBB distribution

Three parameters to be estimated which are a,b and c. They have to represent a unique combination as a whole.

* a,b,c in the domain region of greater than zero.

```{r printing estimated parameters for McGBB Distribution, echo=FALSE}
#estimating parameters for McGBB Distribution of Male_children data
Est_para<-suppressWarnings(mle2(EstMLEMcGBB,
                           data=list(x=Male_Children$No_of_Males,freq=Male_Children$freq),
                           start=list(a=1,b=10,c=20)))
cat("Estimated parameter a for Male_children data=",coef(Est_para)[1],"\n")
cat("Estimated parameter b for Male_children data=",coef(Est_para)[2], "\n")
cat("Estimated parameter c for Male_children data=",coef(Est_para)[3], "\n")
```
